18th International Satisfiability Modulo Theories Competition (SMT-COMP 2023): Rules and Procedures

### 2024-04-02 rules/rules24.tex
First draft of the rules for 2024.

@@

\author{%
Fran\c{c}ois Bobot\\
CEA List \\
France \\
{\small\href{mailto:Francois.BOBOT@cea.fr}{\texttt{https://github.com/bobot}}} \\
\and
Martin Bromberger\\
Max Planck Institute for Informatics\\
Saarland Informatics Campus\\
Germany\\
{\small\href{mailto:mbromber@mpi-inf.mpg.de}{\texttt{mbromber@mpi-inf.mpg.de}}}\\
\and
Jochen Hoenicke\\
Certora\\
Germany\\
{\small\href{mailto:jochen@certora.com}{\texttt{jochen@certora.com}}}
}

@@

As in previous years, we have revised the rules slightly.  Some rule
changes in 2023 are designed to improve how logics are organized into
divisions. We have also added more divisions to the \mvaltrack and have defined a
new \prooftrack.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf System descriptions.}  \emergencystretch.5cm The system description 
  for a participating solver must be submitted by the deadline for the preliminary 
  solver versions. In the case a system description contains not all necessary information,
  participants will be asked to submit an updated version of their 
  system description as part of the submission of the final version of their solver.
%
  \rationale{We need this time to check whether the system descriptions contain all
  necessary information and to give participants the chance to fix their system 
  descriptions if necessary.}

  \item {\bf System descriptions.}  \emergencystretch.5cm Some formulations 
  in the requirements for the system descriptions have been updated from 
  \emph{should} to \textbf{must} to make it clearer that these requirements 
  are mandatory and not just recommended/optional.
%
  \rationale{We did this to ensure that derived tools are more easily recognizable and 
  that it is always clear in which way wrapper tools and derived tools improve 
  on the original solvers. 
  Our changes follow the key words for use in RFCs to indicate requirement levels 
  as proposed by \url{https://www.ietf.org/rfc/rfc2119.txt}.}

  \item {\bf \mvaltrack.}  \emergencystretch.5cm Last year's
    competition introduced the experimental divisions QF\_FPArith
    (excluding the logics QF\_ABVFP, QF\_AUFBVFP, and QF\_ABVFPLRA).
  %
  This year these divisions will be officially counted and no longer be experimental.
  %
  In addition, we add this year new experimental divisions for
  arrays, non-linear arithmetic and for datatypes.  These include the divisions
  QF\_NonLinearIntArith, QF\_NonLinearRealArith,
  QF\_Equality+NonLinearArithm,
  and QF\_Datatypes.
%
  \rationale{Last year's divisions ran successfully, so there
    is no reason to keep them experimental.
    %
    For the new divisions, as before, given the inconsistencies across different
    model producing solvers, we proceed in an experimental fashion to push for
    model standardization.}

\item {\bf \prooftrack.}
  Last year we added this as experimental track. We will continue with
  this track in more or less the same format.  We will reduce the
  number of benchmarks, though.
%  
  \rationale{We received several submissions that showed that there is
    interest in a proof track.

    We are still far away from a common proof format, so changing the
    track to a competition track or making it non-experimental is not
    feasible at this time.

    The number of benchmarks and the size of the produced proofs
    turned out to be a major stress to the StarExec infrastructure,
    while not giving only little insight.  Therefore, we will reduce
    the number of benchmarks in this track.
  }

\item {\bf Wrapper Script for \prooftrack.}
   The wrapper script is changed to remove very large proofs and
   to cut very large error/debug output.
%
   \rationale{Last Year, the \prooftrack produced 800~GB of output, most
   of that stemming from a few benchmark/solver pairs. These cannot be
   removed without losing the whole proof track job.   It also makes
   downloading and sharing the result much harder.}

\item {\bf Scoring.}  \emergencystretch.5cm We adjusted the wall-clock and 
CPU time scores so only the benchmarks that are correctly solved within the 
time limit $T$ are counted.
%
  \rationale{We noticed that in the previous scoring unintended results 
  did occur because timeouts were scored worse than aborts and could 
  overshadow the times of correctly solved problems. For instance, 
  solver A that never aborts could lose to solver B that frequently aborts 
  although A solved exactly the same problems B could solve and A was always 
  faster on the solved problems than B.}
\end{itemize}

@@

\header{SMT Solver Submission.}
%
An entrant to SMT-COMP is a solver submitted by its authors using
the StarExec (\url{http://www.starexec.org}) service, or, for
\paralleltrack{} and \cloudtrack{}, otherwise communicated to the
organisers.

\header{Solver execution.}
%
The StarExec execution
service enables members of the SMT research community to run solvers
on jobs consisting of benchmarks from the SMT-LIB benchmark library.
Jobs are run on a shared computer cluster.  The execution service is
provided free of charge, but requires registration to create a
login account.  Registered users may then upload solvers to
run, or may run public solvers already uploaded to the service.
Information about how to configure and upload a solver is contained in
the StarExec user guide,
\url{https://wiki.uiowa.edu/display/stardev/User+Guide}.

\header{Participation in the Competition.}
%
For participation in SMT-COMP, a solver must be uploaded to StarExec
and made publicly available, or communicated separately to the
organisers for the \cloudtrack{} and \paralleltrack{}.  StarExec
supports solver configurations.  You can have at most one configuration
for each track (named after the track) and a default configuration used
for tracks without a track-specific configuration.
Moreover, the organizers must be informed of the solver's presence
\emph{and the tracks and divisions which it enters} via the web form
at
\begin{center}
  \url{https://forms.gle/xhUDQaTynbS3mccE7}
\end{center}
For each track the submission must specify the logics which are
supported by the solver.  The solver will enter all divisions where it
supports at least one logic.  A submission \textbf{must} also include a
link to the \emph{system description} (see below) and a \emph{32-bit
  unsigned integer}.  These integer numbers, collected from all
submissions, are used to seed competition tools.

@@

SMT-COMP entrants must be submitted via StarExec (solvers), or
communicated separately to the organisers for the \paralleltrack{} and
\cloudtrack{}, \emph{and}
the above web form (accompanying information) until the end of
{\bf May~13, 2023} anywhere on earth.
After this date \emph{no new entrants} will be accepted.
However, updates to existing entrants on StarExec or \paralleltrack{}
and \cloudtrack{}
will be accepted until the end of {\bf May~27, 2023} anywhere on earth.

We strongly encourage participants to use this grace period
\emph{only} for the purpose of fixing any bugs that may be discovered,
and not for adding new features, as there may be no opportunity to do
extensive testing using StarExec or other means after the initial deadline.

The solver versions that are present on StarExec or communicated
otherwise to the organisers for \paralleltrack{} and \cloudtrack{} at the conclusion of
the grace period will be the ones used for the competition.  Versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the day after the final solver
deadline.

@@

\header{Tools.} \label{tools}
The competition uses a number of tools/scripts to run the competition. In the
following, we briefly describe these tools. Unless stated otherwise, these
tools are found at \url{https://github.com/SMT-COMP/smt-comp/tree/master/tools}.
\begin{itemize}
  \item \textbf{Benchmark Selection.} We use a script to implement the
    benchmark selection policy described on page~\pageref{benchmark-selection}.
    It takes a seed for the random benchmark selection. The same seed is used
    for all tools requiring randomisation.
  \item \textbf{Scrambler.} This tool is used to scramble benchmarks during the
    competition to ensure that tools do not rely on syntactic features to
    identify benchmarks. The scrambler can be found at
    \url{https://github.com/SMT-COMP/scrambler}.
  \item \textbf{Trace Executor.} This tool is used in the \inctrack to emulate
    an on-line interaction between an SMT solver and a client application and
    is available at \url{https://github.com/SMT-COMP/trace-executor}
  \item \textbf{Post-Processors.} These are used by StarExec to translate the
    output of tools to the format required for scoring. All post-processors (per
    track) are available at \url{https://github.com/SMT-COMP/postprocessors}.
  \item \textbf{Scoring.} We use a script to implement the scoring computation
    described on in Section~\ref{sec:scoring}. It also includes the scoring
    computations used in previous competitions (since 2015).
\end{itemize}

@@

The StarExec service also limits the memory consumption of the solver
processes.  We expect the memory limit per solver/benchmark pair to be
on the order of 60\,GB.  The values of both the time limit and the
memory limit are available to a solver process through environment
variables.  See the StarExec user guide for more information.

@@

\header{Persistent State.}
%
Solvers may create and write to files and directories during the
course of an execution, but they must not read such files back during
later executions.  Each solver is executed with a temporary directory
as its current working directory.  Any generated files should be
produced there (and not, say, in the system's \texttt{/tmp}
directory).  The StarExec system sets a limit on the amount of disk
storage permitted---typically 20\,GB.  See the StarExec user guide for
more information.  The temporary directory is deleted after the job is
complete.  Solvers must not attempt to communicate with other
machines, e.g., over the network.

@@

\header{Wrapper and Post-Processor.}
Solvers in this track will be wrapped in the competition to include the script
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/starexec_run_proof}},
which calls the solver and checker script.
Participants are encouraged to wrap their solvers themselves to test
their submission.
The track will use
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/process}}
as a post-processor
to validate and accumulate the results.

@@

\subsubsection{\inctrack}
%
An application benchmark may contain multiple \akey{check-sat}
commands.  Solvers may partially solve the benchmark before timing
out.  The benchmark is run by the trace executor, measuring the total
time (summed over all individual commands) taken by the solver to
respond to commands.\footnote{Times measured by StarExec may include
  time spent in the trace executor.  We expect that this time will
  likely be insignificant compared to time spent in the solver, and
  nearly constant across solvers.}  Most time will likely be spent in
response to \akey{check-sat} commands, but \akey{assert}, \akey{push}
or \akey{pop} commands might also entail a reasonable amount of
processing.  For the \inctrack, we have
\begin{itemize}
\item $e=1$ and $n=0$ if the solver returns an incorrect result for any
  \akey{check-sat} command within the time limit,
\item otherwise, $e=0$ and $n$ is the number of correct results for
  \akey{check-sat} commands returned by the solver before the time
  limit is reached.
\end{itemize}

@@

Many others have contributed benchmarks, effort, and feedback.  Clark Barrett,
Pascal Fontaine, Aina Niemetz and Mathias Preiner are maintaining the SMT-LIB
benchmark library.
The competition uses the
\href{https://www.starexec.org/}{StarExec} service, which is hosted at
the \href{http://www.cs.uiowa.edu/}{University of Iowa}.  Aaron Stump
is providing essential StarExec support.

\header{Disclosure.}
Fran\c{c}ois is part of the developing team of the SMT solver COLIBRI~\cite{colibri}.
%
Martin is part of the developing team of the SMT solver
SPASS-SATT~\cite{spass-satt}.
%
Jochen is part of the developing team of the SMT solver
SMTInterpol~\cite{smtinterpol}.

>>>>

\newcommand{\executionservice}{execution service\xspace}

@@

\author{%
Fran\c{c}ois Bobot\\
CEA List \\
France \\
{\small\href{mailto:Francois.BOBOT@cea.fr}{\texttt{https://github.com/bobot}}} \\
\and
Martin Bromberger\\
Max Planck Institute for Informatics\\
Saarland Informatics Campus\\
Germany\\
{\small\href{mailto:mbromber@mpi-inf.mpg.de}{\texttt{mbromber@mpi-inf.mpg.de}}}\\
\and
Martin Jon\'{a}\v{s}\\
Masaryk University\\
Czechia\\
{\small\href{mailto:martin.jonas@mail.muni.cz}{\texttt{martin.jonas@mail.muni.cz}}}
}

@@

As in previous years, we have revised the rules slightly.  Some rule
changes in 2023 are designed to improve how logics are organized into
divisions. We have also added more divisions to the \mvaltrack and have defined a
new \prooftrack.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf Submission process.}  \emergencystretch.5cm \todo{Details}.
%
  \rationale{\todo{Details}}

  \item {\bf Move from StarExec to BenchExec.}  \emergencystretch.5cm \todo{Details}
%
  \rationale{\todo{Details}}
\end{itemize}

@@

\header{SMT Solver Submission.}
%
An entrant to SMT-COMP is a solver submitted by its authors via
a push request to the SMT-COMP GitHub repository (\url{https://github.com/SMT-COMP/smt-comp.github.io/tree/master/submissions}), or, for
\paralleltrack{} and \cloudtrack{}, otherwise communicated to the
organisers. The final solver version needs to be uploaded as a Zenodo (https://zenodo.org/ <https://zenodo.org/>) submission 
that must contain a docker file that compiles the solver.

\header{Solver execution.}
%
BenchExec (\url{https://github.com/sosy-lab/benchexec}) is a framework for reliable benchmarking and resource 
measurement developed by LMU's Software and Computational Systems Lab (SoSy-Lab \url{https://www.sosy-lab.org/}).
The framework itself can be downloaded and used locally by anyone. 
The competition itself will be executed on a BenchExec cluster owned by SoSy-Lab, 
who are kind enough to support our competition with their computing power. 
\todo{Is there a way for the users to run tests directly on the SoSy-Lab BenchExec cluster?}
%
%
%service enables members of the SMT research community to run solvers
%on jobs consisting of benchmarks from the SMT-LIB benchmark library.
%Jobs are run on a shared computer cluster.  The execution service is
%provided free of charge, but requires registration to create a
%login account.  Registered users may then upload solvers to
%run, or may run public solvers already uploaded to the service.
%Information about how to configure and upload a solver is contained in
%the StarExec user guide,
%\url{https://wiki.uiowa.edu/display/stardev/User+Guide}.

\header{Participation in the Competition.}
%
For participation in SMT-COMP, the organizers must be informed of the solver's presence
\emph{and the tracks and divisions which it enters} via a push request 
to the SMT-COMP GitHub repository
\begin{center}
  \url{https://github.com/SMT-COMP/smt-comp.github.io/tree/master/submissions}
\end{center}
The easiest way to create this push request is via the web-interface available under 
the above URL. The actual solver must be submitted via a link to a publicly available 
docker file that compiles the solver. For the \cloudtrack{} and \paralleltrack{},
it must be communicated separately to the organisers by email to 
\href{mailto:aws-smtcomp-2024@googlegroups.com}{aws-smtcomp-2024@googlegroups.com}. 
Note that independent of the tracks, the final solver version must be uploaded as a Zenodo 
(https://zenodo.org/ <https://zenodo.org/>) submission.
Moreover, for each track, the submission must specify the logics which are
supported by the solver. The solver will enter all divisions where it
supports at least one logic. It is possible to submit specific configurations 
of command line  options per track/logic/division as well as a default configuration 
of command line options for all tracks/logics/divisions without their own configuration.
A submission \textbf{must} also include a link to the \emph{system description} (see below) 
and a \emph{32-bit unsigned integer}. These integer numbers, collected from all
submissions, are used to seed competition tools.

@@

SMT-COMP entrants must be submitted via a push request as explained above and,
in the case of the \paralleltrack{} and \cloudtrack{}, 
communicated separately to the organisers via email until the end of
{\bf May~27, 2024} anywhere on earth.
After this date \emph{no new entrants} will be accepted.
However, updates to existing entrants via push requests or,
in the case of the \paralleltrack{} and \cloudtrack{}, via email,
will be accepted until the end of {\bf June~8, 2024} anywhere on earth.

We strongly encourage participants to use this grace period
\emph{only} for the purpose of fixing any bugs that may be discovered,
and not for adding new features, as there may be no opportunity to do
extensive testing using the \executionservice or other means after the initial deadline.

The last solver links pushed to the repository or communicated
via email to the organisers for \paralleltrack{} and \cloudtrack{} at the conclusion of
the grace period will be the ones used for the competition.  Versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the day after the final solver
deadline.

@@

\header{Tools.} \label{tools}
The competition uses a number of tools/scripts to run the competition. In the
following, we briefly describe these tools. Unless stated otherwise, these
tools are found at \url{https://github.com/SMT-COMP/smt-comp/tree/master/tools}.
\begin{itemize}
  \item \textbf{Benchmark Selection.} We use a script to implement the
    benchmark selection policy described on page~\pageref{benchmark-selection}.
    It takes a seed for the random benchmark selection. The same seed is used
    for all tools requiring randomisation.
  \item \textbf{Scrambler.} This tool is used to scramble benchmarks during the
    competition to ensure that tools do not rely on syntactic features to
    identify benchmarks. The scrambler can be found at
    \url{https://github.com/SMT-COMP/scrambler}.
  \item \textbf{Trace Executor.} This tool is used in the \inctrack to emulate
    an on-line interaction between an SMT solver and a client application and
    is available at \url{https://github.com/SMT-COMP/trace-executor}
  \item \textbf{Post-Processors.} These are used by the \executionservice to translate the
    output of tools to the format required for scoring. All post-processors (per
    track) are available at \url{https://github.com/SMT-COMP/postprocessors}.
  \item \textbf{Scoring.} We use a script to implement the scoring computation
    described on in Section~\ref{sec:scoring}. It also includes the scoring
    computations used in previous competitions (since 2015).
\end{itemize}

@@

The StarExec service also limits the memory consumption of the solver
processes.  We expect the memory limit per solver/benchmark pair to be
on the order of 60\,GB.  The values of both the time limit and the
memory limit are available to a solver process through environment
variables.  See the StarExec user guide for more information.
\todo{What are the limits for BenchExec? Do we set limits ourselves?}

@@

\header{Persistent State.}
%
Solvers may create and write to files and directories during the
course of an execution, but they must not read such files back during
later executions.  Each solver is executed with a temporary directory
as its current working directory.  Any generated files should be
produced there (and not, say, in the system's \texttt{/tmp}
directory).  The StarExec system sets a limit on the amount of disk
storage permitted---typically 20\,GB.  See the StarExec user guide for
more information.  The temporary directory is deleted after the job is
complete.  Solvers must not attempt to communicate with other
machines, e.g., over the network.
\todo{What are the limits for BenchExec? Do we set limits ourselves?}

@@

\header{Wrapper and Post-Processor.}
Solvers in this track will be wrapped in the competition to include the script
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/starexec_run_proof}},
which calls the solver and checker script.
Participants are encouraged to wrap their solvers themselves to test
their submission.
The track will use
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/process}}
as a post-processor
to validate and accumulate the results.
\todo{We need a new link for the corresponding benchexec script.}

@@

\subsubsection{\inctrack}
%
An application benchmark may contain multiple \akey{check-sat}
commands.  Solvers may partially solve the benchmark before timing
out.  The benchmark is run by the trace executor, measuring the total
time (summed over all individual commands) taken by the solver to
respond to commands.\footnote{Times measured by the \executionservice may include
  time spent in the trace executor.  We expect that this time will
  likely be insignificant compared to time spent in the solver, and
  nearly constant across solvers.}  Most time will likely be spent in
response to \akey{check-sat} commands, but \akey{assert}, \akey{push}
or \akey{pop} commands might also entail a reasonable amount of
processing.  For the \inctrack, we have
\begin{itemize}
\item $e=1$ and $n=0$ if the solver returns an incorrect result for any
  \akey{check-sat} command within the time limit,
\item otherwise, $e=0$ and $n$ is the number of correct results for
  \akey{check-sat} commands returned by the solver before the time
  limit is reached.
\end{itemize}

@@

Many others have contributed benchmarks, effort, and feedback.  Clark Barrett,
Pascal Fontaine, Aina Niemetz and Mathias Preiner are maintaining the SMT-LIB
benchmark library.
The competition uses a \href{https://github.com/sosy-lab/benchexec}{BenchExec} cluster 
owned by LMU's Software and Computational Systems Lab (SoSy-Lab https://www.sosy-lab.org/ <https://www.sosy-lab.org/>), 
who are kind enough to support our competition with their computing power. Dirk Beyer 
is providing essential BenchExec support.

\header{Disclosure.}
Fran\c{c}ois Bobot is part of the developing team of the SMT solver COLIBRI~\cite{colibri}.
%
Martin Bromberger is part of the developing team of the SMT solver
SPASS-SATT~\cite{spass-satt}.
%
Martin Jon\'{a}\v{s} is part of the developing team of \todo{Martin J. please fill this out}.
